---
title: "lab_10"
author: "Derek Willis"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries, plus some of the map-specific ones from the pre_lab. You'll also be fetching Census data in Q3. Read the questions carefully!

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Load libraries
library(tidyverse)
library(sf)
library(janitor)
library(tigris)
library(lubridate)
library(janitor)
```

Let's explore Maryland elections, power outages and electric vehicle data and make some maps! 

## Questions 

**Q1.** Maryland voters made history in November by electing the state's first two African-American statewide office holders: Wes Moore as governor and Anthony Brown as attorney general. You want to compare the two's electoral performances. 

Using the `md_statewide_county.csv` file in the data folder:

#1
calculate the total number of votes for both the governor's race and the attorney general's race 

```{r}
md_statewide_county <- read_csv("/Volumes/ext4GB/camacho-alisha-jour-772/lab_10/data/md_statewide_county.csv")

md_statewide_county

```
#gov race counts by county + total sum

```{r}
#install.packages("kableExtra")

md_race_counts <- md_statewide_county %>% 
 rowwise() %>% 
  mutate(gov_total = sum(Cox, Moore, Lashar, Wallace, Harding, Sefcik, GovWritein)) %>% 
  mutate(ag_total = sum(Peroutka, Brown, AgWritein)) %>% 
adorn_totals("row", "md_gov") %>% 
as.data.frame()

md_race_counts 

```

#2

then percentages for both Moore and Brown _based on the total of their respective contests_. Then calculate the difference between their percentages.


```{r}
moore_v_brown <- md_race_counts %>% 
  rowwise() |> 
  mutate(p_moore = Moore/sum(Cox, Moore, Lashar, Wallace, Harding, Sefcik, GovWritein)*100) |> 
  rowwise() |> 
  mutate(p_brown = Brown/sum(Peroutka, Brown, AgWritein)*100) |> 
  mutate(difference = p_moore-p_brown) |> 
  arrange(difference)
          
moore_v_brown2 <- moore_v_brown |> 
  select(GEOID, County, difference) |> 
  arrange(difference)

moore_v_brown2
moore_v_brown
```

#4
Get the geometry for Maryland's counties using tigris 


```{r}
#get counties
md_counties <- counties() %>%
  filter(STATEFP == "24") %>% 
  rename(County = "NAME")

md_counties  
```
#5

and, joining that dataframe to your results dataframe

```{r}
#join data

counties_moore_brown <- md_counties %>% 
  left_join(moore_v_brown, by = "GEOID")

counties_moore_brown

```

#6
make a county-level map of the differences between Moore and Brown's percentages. What is the story here, and what is the county with the most interesting results?

**A1.** 

I suppose I'd be curious to learn why Worcester, Carroll, Kent and Queen Anne's had the greatest difference in votes between Moore and Brown. 
 

```{r}
ggplot() +
  geom_sf(data = counties_moore_brown, aes(fill=difference)) + 
  theme_minimal() +
  scale_fill_viridis_b(limits = c(-2, 8))

```



**Q2.** Make a county-level map to explore the number of customers impacted by power outages in Maryland. 

#1
Load the `county_outages_march2023.csv` file in the data folder 

```{r}
county_outages <- read.csv("/Volumes/ext4GB/camacho-alisha-jour-772/lab_10/data/county_outages_march2023.csv")

county_outages <- county_outages |> 
  mutate(outage_customers = as.numeric(outage_customers)) |> 
  mutate(total_customers = as.numeric(total_customers))

county_outages

```

#2
and calculate a rate of customers with outages per 10,000 total customers for each county.

```{r}
outage_rate <- county_outages |> 
  mutate(rate = (outage_customers/total_customers)*10000) |> 
  rename(County = "county") |> 
  arrange(desc(rate))

outage_rate

```

#3
Then join that dataframe to your Maryland counties dataframe you made in Q1 

```{r}
outage_rate_county<- md_counties |> 
left_join(outage_rate, by = "County")

outage_rate_county
```

#4 
and create a county-level map of the results; you may need to tweak the fill scale to make the distinctions clearer.

```{r}

ggplot() +
  geom_sf(data = outage_rate_county, aes(fill = rate)) + 
  theme_minimal() + 
  scale_fill_viridis_b(limits = c(0, 400))
  #scale_fill_viridis_b(trans = "log")

```

What's the potential story here? Given access to other months' outage data, what would your next steps be in reporting a story?

**A2.** 

I would be looking into why Talbot County had so many outages in March. My next step would be to compare outages across counties for months with data available. Ideally, I'd like to get a year of data to quickly determine whether there are any trends over time. I'd also look into local news/do research on any information available aboout the outages that month, and then figure out the right office to call in Talbot county with informed questions prepared. 



**Q3.** You've been assigned a story about the use of electric vehicles in Maryland and you want to make a map of the prevalence of electric vehicles in each county. Load the `ev_2020_2023.csv` file, which has the number of electric and hybrid vehicles per county from July 2020 and March 2023 

#1

```{r}
ev_county <- read_csv("/Volumes/ext4GB/camacho-alisha-jour-772/lab_10/data/ev_2020_2023.csv")

ev_county

```

#2
calculate a percentage change for each county (remember, that's `new-old/old`).

```{r}
ev_county_roc <- ev_county |> 
  mutate(roc = (march_2023-july_2020)/july_2020*100) |> 
  rename(County = "county") |> 
  arrange(roc)

ev_county_roc

```

#3
Using the Maryland counties dataframe you made for Q1, make a county map showing the percentage change.

```{r}
#join
ev_county_roc_join <- md_counties |> 
left_join(ev_county_roc, by = "County")

ev_county_roc_join
```

```{r}
#map 
ggplot() +
  geom_sf(data = ev_county_roc_join, aes(fill = roc)) + 
  theme_minimal() +
  #scale_fill_viridis_b(limits = c(100, 450))
  scale_fill_viridis_b(trans = "log")
  
```

What counties stand out to you? What questions do you have about the changes you're seeing?

#A3.1
Caroline and Dorchester counties stand out to me, with the greatest and lowest rates of change. I also find it interesting that they border each other. My first question would be what is going on in these two counties to have different rates compared to the rest of the state. 

##

Next, use tidycensus to retrieve county population data for Maryland from 2021 (you did this in the pre_lab) 

```{r}
library(tidycensus)

md_county_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              year = 2021,
              state = "MD")

md_county_population

```

and join it to your vehicles dataframe. HINT: you may not need to join it on the county name, and you should already have the geometry!

```{r}
ev_county_roc_join_pop <- ev_county_roc_join |> 
left_join(md_county_population, by = "GEOID")

ev_county_roc_join_pop

```

Then, calculate a per capita rate of electric & hybrid vehicles using the March 2023 figure and the population estimate. You will need to choose the "per capita" value.

```{r}

ev_county_roc_pop <- ev_county_roc_join_pop |> 
  select(County, march_2023, estimate) |> 
  rename(population = "estimate") |> 
  mutate(per_capita = (march_2023/population)*10000)

ev_county_roc_pop

```

Finally, make a second county map showing the per capita rate. What counties stand out on this map? How is it different from the first one you made?

**A3.** 

Howard and Montgomery counties stand out as having the most ev/hybrid vehicles in Maryland per person. Compared to the previous map, Caroline County's increase in EV/cars is not significant when population is taken into account. 

```{r}

ggplot() +
  geom_sf(data = ev_county_roc_pop, aes(fill = per_capita)) + 
  theme_minimal() +
  #scale_fill_viridis_b(limits = c(0, 250))
  scale_fill_viridis_b()

```

